Geometrization of characters of tori over non-Archimedean local fields

### 2013-08-13 shanghai.tex
Adding definitions . . . .

% !TEX encoding = UTF-8 Unicode
\documentclass[11pt]{amsart}
\title[Geometrization of characters of tori]{Geometrization of characters of tori over non-Archimedean local fields}
\usepackage[british]{babel}
\usepackage{datetime}
\date{\today}
\author{Clifton Cunningham}
\address{University of Calgary}
\email{cunning@math.ucalgary.ca}
\author{David Roe}
\address{Pacific Institute for the Mathematical Sciences at the University of Calgary}
\email{roed.math@gmail.com}
\usepackage[utf8]{inputenc}
\renewcommand{\baselinestretch}{1.2}
%\usepackage[notcite,color]{showkeys}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage[shortalphabetic]{amsrefs}
\usepackage[alphabetic]{amsrefs}
\renewcommand\MR{\relax}
\usepackage{xypic}
\usepackage{textcomp}
\usepackage{mathrsfs}
\usepackage{yfonts}
\newcommand{\mathswab}[1]{\operatorname{\textswab{#1}}}
\usepackage[T1]{fontenc}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,calc,matrix}
\usepackage{tikz-cd}
\usepackage{manfnt}
%\usepackage{makeidx}

@@

%\begin{proposition}\label{tensor}
The category $\QC(G)$ of quasicharacter sheaves on $G$ is a rigid symmetric monoidal category \Clifton{I still haven't found a good reference for this!} under
$\cs{L} \otimes \cs{L'} \ceq (\gcs{L}\otimes\gcs{L'}, \mu\otimes\mu', \phi\otimes \phi')$
with duals given by applying the functor 
$\sheafHom(\ - \ ,\EE)$ (sheaf hom) to
conditions \ref{CS.0}, \ref{CS.1} and \ref{CS.2}.
%\end{proposition}
(The category of characters sheaves on $G$ is not abelian, so $\QC(G)$
is not a tensor category in the sense of \cite{Deligne:tensorielles}.)
%\begin{proof}\end{proof}
%
This rigid monoidal category structure for $\QC(G)$ puts a group
structure on the set $\QCiso{G}$ of isomorphism classes in
$\QC(G)$. Our main result regarding quasicharacter sheaf sheaves on $G$ is the determination of $\QCiso{G}$ under the hypothesis that the component group scheme for $G$ is geometrically finitely generated; see Theorem~\ref{thm:snake}. To do this, we use the trace of Frobenius.

\begin{definition}
Let $(\gcs{L},\phi)$ be a Weil sheaf on $G$. on constructible $\ell$-adic sheaf on $\bG$. Every $g\in G(\Fq)$
determines a point $\bg$ on $\bG$ fixed by $\Frob{G}$ and
therefore an automorphism $\phi_{\bg}$ of $\gcs{L}_{\bg}$. The \emph{trace of Frobenius} of $(\gcs{L},\phi)$ at $g\in G(\Fq)$ is 
\[
\trFrob{(\gcs{L},\phi)}(g) \ceq \trace(\phi_{\bg}, \gcs{L}_{\bg}).
\]
If $\cs{L} = (\gcs{L},\mu,\phi)$ is a geochar, then we will abbreviate write $\trFrob{\cs{L}}$ for $\trFrob{(\gcs{L},\phi)}$
\end{definition}

The isomorphism $\mu : \bm^* \gcs{L} \to \LxL$
guarantees that the function $\trFrob{\cs{L}} : G(\Fq)\to \EEx$,
defined by $g \mapsto \phi_{\bg}$, is a group homomorphism;
we refer to $\trFrob{\cs{L}}$ as the \cdef{trace of Frobenius of $\cs{L}$}.
Although though no trace is involved in the definition, the notion extends 
to any constructible $\ell$-adic Weil sheaf $(\gcs{L},\phi)$ on $G$, in which context,
$\trFrob{(\gcs{L},\phi)} : G(\Fq)\to \EE$ is defined by $\trFrob{(\gcs{L},\phi)}(g) \ceq \trace \phi_{\bg}$.

It is clear that $\trFrob{\cs{L}\otimes \cs{L'}} = \trFrob{\cs{L}} \cdot \trFrob{\cs{L'}}$
(pointwise multiplication of functions) and that $\trFrob{\cs{L}} = \trFrob{\cs{L'}}$
if $\cs{L} \iso \cs{L'}$ in $\QC(G)$. In this way we obtain a group homomorphism
\[
\cdef{\trFrob{G}} : \QCiso{G} \longrightarrow G(\Fq)^*,%\Hom_\text{grp}(G(\Fq),\EEx),
\qquad \text{defined by}\qquad \cs{L} \mapsto \trFrob{\cs{L}},
\]
from the Picard group of quasicharacter sheaves on $G$ to the character group of $G(\Fq)$.

Our main result on the geometrization of admissible characters of
$p$-adic tori [insert internal reference here] follows from a general
fact established later in this paper: if the geometric component group
of $G$ is finitely-generated then the trace of Frobenius defines an
isomorphism from the Picard group of quasicharacter sheaves $G$ to the
group of characters of $G(\Fq)$ (here, as throughout the paper, $G$ is
a commutative group scheme, smooth and locally of finite type over $\Fq$).

\Clifton{Address functoriality here? Yes, but put proof at the end of the next section.}

In this paper we $\QCiso{G} \iso G(\Fq)^*$ when the geometric component group of $G$ is finitely generated; see Theorem~\ref{thm:snake}. We also shows that this isomorphism is canonical; see

>>>>

% !TEX encoding = UTF-8 Unicode
\documentclass[11pt]{amsart}
\title[Geometrization of characters of tori]{Geometrization of characters of tori over non-Archimedean local fields}
\usepackage[british]{babel}
\usepackage{datetime}
\date{\today}
\author{Clifton Cunningham}
\address{University of Calgary}
\email{cunning@math.ucalgary.ca}
\author{David Roe}
\address{Pacific Institute for the Mathematical Sciences at the University of Calgary}
\email{roed.math@gmail.com}
\usepackage[utf8]{inputenc}
\renewcommand{\baselinestretch}{1.2}
\usepackage[notcite,color]{showkeys}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage[shortalphabetic]{amsrefs}
\usepackage[alphabetic]{amsrefs}
\renewcommand\MR{\relax}
\usepackage{xypic}
\usepackage{textcomp}
\usepackage{mathrsfs}
\usepackage{yfonts}
\newcommand{\mathswab}[1]{\operatorname{\textswab{#1}}}
\usepackage[T1]{fontenc}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,calc,matrix}
\usepackage{tikz-cd}
\usepackage{manfnt}
%\usepackage{makeidx}

@@

%\begin{proposition}\label{tensor}
The category $\QC(G)$ of quasicharacter sheaves on $G$ is a rigid symmetric monoidal category \Clifton{I still haven't found a good reference for this!} under
$\cs{L} \otimes \cs{L'} \ceq (\gcs{L}\otimes\gcs{L'}, \mu\otimes\mu', \phi\otimes \phi')$
with duals given by applying the functor 
$\sheafHom(\ - \ ,\EE)$ (sheaf hom) to
conditions \ref{CS.0}, \ref{CS.1} and \ref{CS.2}.
%\end{proposition}
(The category of characters sheaves on $G$ is not abelian, so $\QC(G)$
is not a tensor category in the sense of \cite{Deligne:tensorielles}.)
%\begin{proof}\end{proof}
%
This rigid monoidal category structure for $\QC(G)$ puts a group
structure on the set $\QCiso{G}$ of isomorphism classes in
$\QC(G)$. Our main result regarding quasicharacter sheaves on $G$ (Theorem~\ref{thm:snake})
determines $\QCiso{G}$ under the hypothesis that the component group scheme for $G$ is geometrically finitely generated; see Theorem~\ref{thm:snake}. To do this, we use the trace of Frobenius.

\begin{definition}
Let $(\gcs{L},\phi)$ be a Weil sheaf on $G$. on constructible $\ell$-adic sheaf on $\bG$. Every $g\in G(\Fq)$
determines a point $\bg$ on $\bG$ fixed by $\Frob{G}$ and
therefore an automorphism $\phi_{\bg}$ of $\gcs{L}_{\bg}$. The \emph{trace of Frobenius} of $(\gcs{L},\phi)$ at $g\in G(\Fq)$ is 
\[
\trFrob{(\gcs{L},\phi)}(g) \ceq \trace(\phi_{\bg}, \gcs{L}_{\bg}).
\]
If $\cs{L} = (\gcs{L},\mu,\phi)$ is a quasicharacter sheaf, 
then we will abbreviate $\trFrob{(\gcs{L},\phi)}$ to $\trFrob{\cs{L}}$.
\end{definition}

The isomorphism $\mu : \bm^* \gcs{L} \to \LxL$ guarantees 
that the function $\trFrob{\cs{L}} : G(\Fq)\to \EEx$ is a group homomorphism.
It is also clear that $\trFrob{\cs{L}\otimes \cs{L'}} = \trFrob{\cs{L}} \cdot \trFrob{\cs{L'}}$
(pointwise multiplication of functions) and that $\trFrob{\cs{L}} = \trFrob{\cs{L'}}$
if $\cs{L} \iso \cs{L'}$ in $\QC(G)$. In this way we obtain a group homomorphism
\[
\cdef{\trFrob{G}} : \QCiso{G} \longrightarrow G(\Fq)^*,
\qquad \text{defined by}\qquad \cs{L} \mapsto \trFrob{\cs{L}}.
\]
This isomorphism is functorial, in the following sense.

\begin{proposition}\label{prop:functorialG}
Trace of Frobenius $\QCiso{G} \to G(\Fq)$ is a natural transformation 
between the following two additive functors from the category of smooth commutative
group schemes locally of finite type over

 to the category of abelian groups: 
\begin{enumerate}
\item[$F_1$:] $T \mapsto \QCiso{\GN{T}}$;
\item[$F_2$:] $T \mapsto \Hom_\text{ad}(T(K),\EEx)$.
\end{enumerate}
\end{proposition}
